Transforming Trash as an Artistic Act

### 2016-01-17 chapter1_final.tex
Relocation, turnitin fixes
Former-commit-id: e8de078896595446f38348c7c970a0b9f22fae30

% Epigraph
\begin{singlespace}
\epigraph{One can even shout out through refuse \ldots}{\hfill---Kurt Schwitters, \textit{Kurt Schwitters}, 1985}
\end{singlespace}

@@

% TODO Summarize
% Change in the production and consumption practices
\quotes{Over the course of the twentieth century, the twin developments of mass production and mass media in the capitalist economies of the Western Countries completed a total transformation of everyday life, reorienting almost every activity toward consumption. Things once locally produced and often handmade were now mass produced and commodified, turning local, artisanal producers into deskilled laborers serving the assembly line} \citep{pye2010trashculture}. As the result of it amount of goods boost and the cost of them significantly reduced. Things once reused again and again, now thrown away because it is more affordable to replace with new one. As the production increased its by-product material waste is also increased. \quotes{The phenomenon of waste comes clearly into focus not merely as a by-product of manufacturing processes, but rather as an integral element in cycles of production and consumption} \citep{pye2010trashculture}.

@@

People do not think what happens after they throw trash away.  It exits from people's life but not from the world. It is stacked to another place. A project conducted by MIT in America explores the journey of trash by placing trackers onto the trash \citep{chen2009mit}. According to their results trash spreads away across the country and this journey takes month. Further it is not limited with the border of America because they are exported to the other countries. In other words American's trash is not only their trash.

The vast amount of discarded items spread through the landfills to oceans. As commodities spread to the every corner of the world, trash do also. It moves from homes to garbage trucks, from streets to landfills and also from landfills to sculptures inside museums. In other words, trash is not only found on the most disgusting and avoided places, but also can be found on the most steril and frequently visited places.

@@

% Different types
Electronic waste, construction and demolition waste, medical waste are some of the various types of waste. They are result of different production and consumption practices. For example nuclear power generation and applications of nuclear fission or nuclear technology produces not only engery but also radioactive waste which is very dangerous for living creatures.

% Different diciplines
Trash attracts the attention of different peoples and disciplines \quotes{ranging from economics to environmental studies, but most particularly by those studying consumerism or material culture} \citep[63]{emgin2012trashion}. For instance for archaeologist dumps near the residential area are one of the significant places to figure out what people consume in the ancient times \citep{rathje1992rubbish}. Moreover scholars argues that there is a relationship between technological and waste. Trash can be seen as a design and technological problem \citep{mcdonough2010cradle}. Further economical aspect of trash can not ignored. \cite[65]{emgin2012trashion} states that there is a huge business based on eco-friendly products and recyclable goods that is increasingly promoted.

% Different people
Various people including garbage collectors and artists interact with trash for different purposes. There are people who collects plastics, scraps and papers from waste bins and landfills. For the ragpickers, collecting others excess and discard is significant source of economical income. They collect and clasify recycle materials to sell them. By collecting and selling them they endure their life. On the other side for some of the people it is not considered as trash. As stated by the editor of the Garbage issue of ReVista, Christmas decorations at Chocó, a poor region on Colombia’s Pacific Coast, are \quotes{all crafted from used tin cans, old newspapers, discarded textiles and found wood objects} \citep{erlick2015editorsletter}. She realized that any of them called the practice as recycling. For them using trash again and again is very natural and it is part of their life. On the contrary for the developed countries trash considered as a thing that must be avoided. There is no place for trash in their life. It can be understood from this anectode that approaches to the garbage are not same for all regions of the world. It can be said that \quotes{perceptions of waste and the value of material are neither static nor universally shared} \citep[xxvi]{zimring2012encyclopedia}.

As noted by \cite{pye2010trashculture} \quotes{at least since the early twentieth century, the concern with discarded things and materials has been a recurring theme in art}. In the beginning of the 20th century objects that are considered as art enter the scope of art making. Using non-art objects in the art opened new dimensions for the language of art and practice art making. This is a revolutionary change in the practice of art making. Production of art and the approach to the art changed dramatically. Picasso and Baraque are the first artists that used non-art object in their works. Later many of artists such as Kurt Schwitters and Joseph Cornell followed them with their collages and assemblages. Later Duchamp challenged the nature of art by using, non-unique, fabricated object to express his ideas. His work changed the way of perceiving art dramatically. With these developments trash has found a place in the art and contemporary artists become interested in trash. They use trash for different purposes and these are explored in detail in the following chapters. Trash find places in many mediums such as sculptures, paintings, assemblages and photographs. Artist express themselves through trash. Portraits are build from trash. Consumption habits of society are criticized through trash. \cite[2]{pye2010trashculture} states that \quotes{trash have influenced, and are also influenced by, cultural products such as films, visual art, museum exhibits and literature}.

@@

%
%****************************************
\section{Structure of the Study}
Theory and artworks are not strictly separated from each other in this thesis. Artworks are explained with the help of theories. This study is structured by many examples of artworks in different part of the thesis. When an argument is presented, immediately an examplory artwork is given and how artists respond this issue is illustrated. Thereby artworks are examined in the theoretical and cultural context. Artworks are used as supportive elements for the stated arguments. However, artworks are discussed with all of their aspects. Only basic information is given and how it is related with this topic explained. By the way, it is better to keep in mind that all works can be analyzed in different contexts in more detail fashion.

@@

In the third chapter, Trash (in) Art, seeks the root of usin discarded materials in the artworks and analyses the sample artworks and artists. As already artworks are mentioned in the different part of the thesis, in this chapter some of them analyzed deeply. Artists' methods and approaches to the subject are stated.

@@

The last chapter is reserved for an overall conclusion of this thesis and further suggestions on the subject and the project.

@@

Artist Alice Bradshaw has published a \quotes{Rubbish Newspaper} which contains these words and shows their place in litreature and artworks. It is based on her extensive researh on this subject.

@@

According to dictionary Merriam-Webster, the word reuse means \quotes{to use again especially in a different way or after reclaiming or reprocessing.} Reusing involves usage of the same product unchanged in form. Reusing lengthens the life of the item or material. The main purpose is to make the item last as long as it can. To reuse is to use something again instead of throwing it away. In this thesis, reuse is explored through the work of the artist.

On the other hand recycle means \quotes{to treat or process (used or waste materials) so as to make suitable for reuse.} In recycling an item, it is processed into a totally new product. Recycling occurs when waste in an unchanged chemical form is used in the same process that created the original product. Crushed glass containers used to make new glass containers, and scrap metal used in foundries can ge given as examples.

@@

Some of the scholars use recycling \citep{cerny1996recycled,herman1998trashformations} for the recreation and transformation of waste. Many scholar used the word recycling when mentioning works use trash and the concept of it. However they are not mentioning the meaning that is decomposing things to the their particles. What they is actually is reusing and combining things, concepts, creating new mixtures. It is hard to say that they use it wrongly, but what they refer is actually upcycling. Creative reuse, inventing new things from discard. Upcycling is a term used by architect and designer William McDonaugh and chemist Michael Braungart and refers to \quotes{the process of converting an industrial nutrient (material) into something of similar or greater value, in its second life} \citep[as cited in][63]{emgin2012trashion}. It is argued that art, in this instance, acts as a tool of transformation and reintroduces into certain orders what was once deemed waste. This theory counters the argument that an object is dead once it is disposed of \citep{emgin2012trashion}. Some of the scholars and artist uses upcycling for their process. This thesis project is about those objects that are recreated from trash through the process of upcycling.

>>>>

\begin{singlespace}
\epigraph{One can even shout out through refuse \ldots}{\hfill---Kurt Schwitters, \textit{Kurt Schwitters}, 1985}
\end{singlespace}

@@

% Change in the production and consumption practices
\cite[11]{banash2013collage} states that:
\begin{quote}
Over the course of the twentieth century, the twin developments of mass production and mass media in the capitalist economies of the Western Countries completed a total transformation of everyday life, reorienting almost every activity toward consumption. Things once locally produced and often handmade were now mass produced and commodified, turning local, artisanal producers into deskilled laborers serving the assembly line.
\end{quote}
As a result of it amount of goods boost and the cost of them significantly reduced. Things once reused again and again, now thrown away because it is more affordable to replace with new one. As the production increased its by-product material waste is also increased. \quotes{The phenomenon of waste comes clearly into focus not merely as a by-product of manufacturing processes, but rather as an integral element in cycles of production and consumption} \citep{pye2010trashculture}.

@@

People do not think what happens after they throw trash away.  It exits from people's life but not from the world. It is stacked to another place. A project conducted by MIT\footnote{Massachusetts Institute of Technology} in America explores the journey of trash by placing trackers onto the trash \citep{chen2009mit}. According to their results trash spreads away across the country and this journey takes month. Further it is not limited with the border of America because they are exported to the other countries. In other words Americans trash is not only their trash.

The vast amount of discarded items spread through the landfills to oceans. As commodities spread to the every corner of the world, trash do also. It moves from homes to garbage trucks, from streets to landfills and also from landfills to sculptures inside museums. In other words, trash is not only found on the most disgusting and avoided places, but also can be found on the most sterile and frequently visited places.

@@

% Different types
Electronic waste, construction and demolition waste, medical waste are some of the various types of waste. They are result of different production and consumption practices. For example nuclear power generation and applications of nuclear fission or nuclear technology produces not only energy but also radioactive waste which is very dangerous for living creatures.

% Different disciplines
Trash attracts the attention of different peoples and disciplines \quotes{ranging from economics to environmental studies, but most particularly by those studying consumerism or material culture} \citep[63]{emgin2012trashion}. For instance for archaeologist dumps near the residential area are one of the significant places to figure out what people consume in the ancient times \citep{rathje1992rubbish}. Moreover scholars argues that there is a relationship between technological and waste. Trash can be seen as a design and technological problem \citep{mcdonough2010cradle}. Further economical aspect of trash can not ignored. \cite[65]{emgin2012trashion} states that there is a huge business based on eco-friendly products and recyclable goods that is increasingly promoted.

% Different people
Various people including garbage collectors and artists interact with trash for different purposes. There are people who collects plastics, scraps and papers from waste bins and landfills. For the rag-pickers, collecting others excess and discard is significant source of economical income. They collect and classify recycle materials to sell them. By collecting and selling them they endure their life. On the other side for some of the people it is not considered as trash. As stated by the editor of the Garbage issue of ReVista, Christmas decorations at Chocó, a poor region on Colombia’s Pacific Coast, are \quotes{all crafted from used tin cans, old newspapers, discarded textiles and found wood objects} \citep{erlick2015editorsletter}. She realized that any of them called the practice as recycling. For them using trash again and again is very natural and it is part of their life. On the contrary for the developed countries trash considered as a thing that must be avoided. There is no place for trash in their life. It can be understood from this anecdote that approaches to the garbage are not same for all regions of the world. It can be said that \quotes{perceptions of waste and the value of material are neither static nor universally shared} \citep[xxvi]{zimring2012encyclopedia}.

As noted by \cite{pye2010trashculture} \quotes{at least since the early twentieth century, the concern with discarded things and materials has been a recurring theme in art}. In the beginning of the 20th century objects that are considered as art enter the scope of art making. Using non-art objects in the art opened new dimensions for the language of art and practice art making. This is a revolutionary change in the practice of art making. Production of art and the approach to the art changed dramatically. Picasso and Braque are the first artists that used non-art object in their works. Later many of artists such as Kurt Schwitters and Joseph Cornell followed them with their collages and assemblages. Later Duchamp challenged the nature of art by using, non-unique, fabricated object to express his ideas. His work changed the way of perceiving art dramatically. With these developments trash has found a place in the art and contemporary artists become interested in trash. They use trash for different purposes and these are explored in detail in the following chapters. Trash find places in many mediums such as sculptures, paintings, assemblages and photographs. Artist express themselves through trash. Portraits are build from trash. Consumption habits of society are criticized through trash. \cite[2]{pye2010trashculture} states that \quotes{trash have influenced, and are also influenced by, cultural products such as films, visual art, museum exhibits and literature}.

@@

%****************************************
\section{Structure of the Study}
Theory and artworks are not strictly separated from each other in this thesis. Artworks are explained with the help of theories. This study is structured by many examples of artworks in different part of the thesis. When an argument is presented, immediately an examplory artwork is given and how artists respond this issue is illustrated. Thereby artworks are examined in the theoretical and cultural context. Artworks are used as supportive elements for the stated arguments. However, artworks are discussed with all of their aspects. Only basic information is given and how it is related with this topic explained. By the way, it is better to keep in mind that all works can be analyzed in different contexts in more detail fashion.

@@

In the third chapter, Trash (in) Art, seeks the root of using discarded materials in the artworks and analyses the sample artworks and artists. As already artworks are mentioned in the different part of the thesis, in this chapter some of them analyzed deeply. Artists' methods and approaches to the subject are stated.

@@

The last chapter is reserved for an overall conclusion and evaluation of this thesis and further suggestions on the subject and the project.

@@

Artist Alice Bradshaw has published a 36 page \quotes{Rubbish Newspaper} which contains these words and illustrates their place in literature and artworks. It is based on her extensive research on this subject.

In this thesis like many scholars I also use these words interchangeably. However mainly in this thesis scope I prefer trash which better signifies dry paper discard.

@@

According to dictionary Merriam-Webster, the word reuse means \quotes{to use again especially in a different way or after reclaiming or reprocessing.} Reusing includes usage of the same product unaltered in structure. Reusing lengthens the life of the item or material. The main purpose is to make the item last as long as it can. To reuse is to use something again instead of throwing it away. In this thesis, reuse is explored through the work of the artist.

On the other hand recycle means \quotes{to treat or process (used or waste materials) so as to make suitable for reuse.} In recycling, an object is processed into a totally new product. Recycling occurs when waste in an unchanged chemical form is used in the same process that created the original product. Crushed glass containers used to make new glass containers, and scrap metal used in foundries can be given as examples.

@@

Some of the scholars use recycling \citep{cerny1996recycled,herman1998trashformations} for the recreation and transformation of waste. Many scholar used the word recycling when mentioning works use trash. However they are not mentioning the meaning that is decomposing things to the their particles. What they actually referring is reusing and combining things, concepts to create new mixtures. It is hard to say that they use it wrongly, but what they refer is actually upcycling. Creative reuse, inventing new things from discard. Upcycling is a term coined by architect and designer William McDonaugh and chemist Michael Braungart and refers to \quotes{the process of converting an industrial nutrient (material) into something of similar or greater value, in its second life} \citep[as cited in][63]{emgin2012trashion}. It is argued that art, in this scope, acts as a tool of transformation and reintroduces into certain orders what was once regarded waste. This approach resists the argument that an object is dead once it is disposed of \citep{emgin2012trashion}. Some of the scholars and artist uses upcycling for their works. Undertaking thesis project is about those objects that are reproduced from trash through the procedure of upcycling.
